---
title: "Getting Started with Multiple Factor Analysis"
author: "Ningning Long, Beatrice Girard, Jordan Prosky, Laura Katz, Xiaoqi Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with Multiple Factor Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(MFA)
```

## MFA

The package `"MFA"` is an implementation for handling multiple data tables given a specific dataset. The main mfa function will create a mfa object according to the dataset, which including information of eigenvalues, singular values, common factor scores, partial factor scores, loadings and partial loadings. Then, after getting the mfa object, there are several related functions and methods that can allow the users to get eigenvalues summary, contributions, Rv coefficient, Lg Coefficient, stability estimation from bootstrap. Also, there is a plot method created for mfa to visualize the compromise of tables, partial factor scores and variable loadings.



## The Dataset

In order to give a tutorial of the MFA package, we choose to use the wine tasting experiment dataset.The raw data is available in the file wines.csv inside the final-project/data/ folder: https://github.com/ucb-stat243/stat243-fall-2016/tree/master/problem-sets/final-project 



## Creating a mfa object

The first step is to create a `"mfa"` object with the function `mfa()`:
```{r}
#get the dataset
wines <- read.csv("https://raw.githubusercontent.com/ucb-stat243/stat243-fall-2016/master/problem-sets/final-project/data/wines.csv", stringsAsFactors = FALSE)

#According to the dataset, creating a list of vectors indicating the sets of variables
sets <- list(2:7, 8:13, 14:19, 20:24, 25:30, 31:35, 36:39, 40:45, 46:50, 51:54)

#The scaling vector
scaling_vec <- apply(subset(wines, select = unlist(sets)), 2, function(x) sqrt(sum((x - mean(x))^2)))

#create the mfa object 
mymfa <- mfa(wines, sets, ncomps = 2, T, scaling_vec)

mymfa

```

Here, `mfa()` creates a mfa object of wines dataset, with specific sets and scaling_vect. Also, we choose ncomps=2 just for convinience. 

You can change the `mfa()` arguments `data`, `sets`, `ncomps`, `center` and `scale`. Please see more descriptions about the arguments in the documentation of `mfa()`.




## Summary of eigenvalues

Once we get the mfa object from the `mfa()` function, we can use it to get the summarizing information about the obtained eigenvalues. The `eigensummary()` method for mfa objects can give a table with singular values, eigenvalues, cumulative, percentage of intertia, cumulative percentage of inertia, for all the extracted components (in this case, our extracted components are the first two components).
```{r}
eigensummary(mymfa)
```



## Contributions

The contributions are a series of descriptive statistics that help us interpret how 1) the observations, 2) the variables, and 3) the tables, contribute to the variability of the extracted dimensions. Once we have the mfa object, we can call the following functions to get contributions: `contrib_obs()`, `contrib_var()`,and `contrib_tables()`
```{r}
#Contribution of an observation to a dimension
contrib_obs1 <- contrib_obs(mymfa)
contrib_obs1


#Contribution of a variable to a dimension
#use data_tables() function to separate data into individual tables (see the documentation of data_tables() for more information)
tables <- data_tables(wines, sets, TRUE, scaling_vec)
#shift the sets vector
sets_new <- lapply(sets, FUN=function(vec){vec-1})
contrib_var1 <- contrib_var(mymfa, sets_new, tables)
contrib_var1


#Contribution of a table to a dimension
contrib_table1 <- contrib_tables(mymfa, sets_new, tables)
contrib_table1

```




## Coefficients to Study the Between-Table Structure

To evaluate the similarity between two tables we use the RV coefficient. The `RV()` function computes this coefficient for two tables. The `RV_table()` function takes a data set, and a list with sets of variables, and which returns a matrix of RV coefficients.
```{r}
#Rv coefficient for two tables
RVtable1_table2 <- RV(tables[[1]], tables[[2]])
RVtable1_table2

#sets of variables
sets_shift1 <- list(1:6, 7:12, 13:18)
#a dataset that comprises three sets or blocks
dataset1 <- data.frame(cbind(tables[[1]], tables[[2]], tables[[3]]))
# get a matrix of RV coefficients between each pair of tables
RV_dataset <- RV_table(dataset1, sets_shift1)
RV_dataset
```



## Lg Coefficient

Similarly, we can also use Lg coefficient in the context of MFA to evaluate the similarity between two tables. The `Lg()` function computes this coefficient for two tables. The `Lg_table()` function takes a data set, and a list with sets of variables, and which returns a matrix of Lg coefficients.
```{r}
#Lg coefficient between two tables
Lgtable1_table2 <- Lg(tables[[1]], tables[[2]])
Lgtable1_table2

#The Lg coefficient matrix between the tables defined by the sets
sets_shift2 <- list(1:6,7:12,13:18,24:29,39:44)
dataset2 <- as.data.frame(tables)
Lg_dataset <- Lg_table(dataset2, sets_shift2)
Lg_dataset
```



## Bootstrap

The bootstrap is a method for mfa objects. The method allows the user to perform bootstrapping in order to estimate the stability of the compromise factor scores. The method takes two inputs (a mfa object and the number of times of bootstrap) and returns a list, containing the mean, standard deviation and bootstrap ratio; Also, it creates plot of bootstrap ratio for each dimension.
```{r}
#do boostrap 1000 times with a mfa object
bootstrap(mymfa, 1000)
```



## Plot Method

We can use the plot method of mfa object to visualize the compromise of tables, partial factor scores, and variable loadings for a given two dimensions
```{r,fig.show='hold'}
##plot for factor scores
plot(mymfa, type = 1)

#plot for partial factor scores and variable loadings with given loading_labels for the first assessor
plot(mymfa, type = 2, X = 1, loading_labels = c("Cat Pee", "Passion Fruit", "Green Pepper", "Mineral", "Smoky", "Citrus"))

#printing all 10 plots for partial factor scores and variable loadings for the ten assessors without given loading_labels
plot(mymfa, type = 3, loading_labels = NULL)

#plot for partial factor scores  for the first assessor
plot(mymfa, type = 4, X = 1) 

#plot for variable loadings for the first assessor
plot(mymfa, type = 5, X = 1) 
```


```{r setup, include=FALSE}
#remove the things in the environment
rm(list = ls())
```

